On Misuses of the Kolmogorov--Smirnov Test for One Sample Goodness-of-Fit

### 2022-04-29 Manuscript/manuscript.tex
textwidth graphs
facet_grid plot, final typos check

The Kolmogorov-Smirnov (KS) statistic is one of the most popular goodness-of-fit 
tests for comparing a sample with a hypothesized parametric distribution.
Let $X_1, ..., X_n$ be a random sample from some continuous distribution. 
The null hypothesis $H_0$ is that the population distribution is $F$.
Let $F_n(\cdot)$ be the empirical cumulative distribution. The KS statistic is
\begin{equation}
  \label{eq:ks_standard}
  D = \sup_x | F_{n}(x) - F(x) |.
\end{equation}
The distribution of $D$ under $H_0$ turned out to be independent of the
distribution $F$ and a table of critical values for $D$ has been constructed
\citep{Massey} for various sample sizes $n$ and significance 
levels $\alpha$. If the value of $d$ exceeds the test's corresponding critical 
value, the null hypothesis is rejected. The KS test is available in popular 
statistical software packages, such as function \texttt{ks.test} in R 
\citep{R, Marsaglia}.

The standard one-sample KS test applies to independent data with a continuous
hypothesized distribution that is completely specified. Often in literature, it 
has been ``discovered" that the test loses its power when applied to dependent 
or discrete data with hypothesized distributions containing estimated parameters 
\citep{Steinskog}.

@@

The standard KS test is not applicable when the hypothesized distribution 
contains fitted parameters. \citet{Steinskog} "discovers" the change in power 
when using estimated parameters and stress caution in using the KS test in such 
ways. \citet{Lilliefors} shows that using the standard table when values of the 
mean and standard deviation are estimated obtains extremely conservative 
results. This is supported by \citet{Capasso}, which concludes that failing to 
re-estimate the parameters may lead to wrong, overly-conservative approximations 
to the distributions of goodness-of-fit test statistics based on the empirical 
distribution function. \citet{Capasso} also notes that the impact of this 
possible mistake may turn out to be dramatic and does not vanish as the sample 
size increases. Remedies are provided by \citet{Genest} and \citet{Babu} in the 
form of bootstrap. \citet{Genest} provides validity for using parametric 
bootstrap with various goodness-of-fit tests. \citet{Babu} details the bootstrap
procedure for goodness-of-fit tests and notes that both parametric and 
non-parametric procedures lead to correct asymptotic levels, however there 
is a correction required for the non-parametric case.

@@

The rest of the paper is organized as follows. Section~\ref{sec:fitted}
investigates the scenario where the hypothesized distribution has unspecified
parameters. Both parametric and nonparametric bootstrap are available to fix the
issue. Section~\ref{sec:dependence} investigates the scenario where the data of 
the empirical distribution is serially dependent. Our bootstrap procedure 
employing copulas is a working solution. Section~\ref{sec:fittedwithdependence}
explores the case where a combination of the first two scenarios occurs. The 
copula procedure is available as a fix here as well. 
Section~\ref{sec:conclusion} concludes with a discussion.

@@

To demonstrate the consequences of this problem, a simulation is performed. A 
random sample $X_1, \ldots, X_n$ is generated from the standard normal 
distribution with sample size $n = 100$. The p-values of $1000$ replicate tests 
are displayed in the Naive plot of Figure~\ref{fig:hist_fitted}. Each KS test 
was performed using fitted parameters, i.e. the hypothesized distribution is 
$N(\bar X, s^2)$ where $\bar X$ is the sample mean and $s^2$ is the sample 
variance. Since the data is generated from a standard normal distribution with 
seemingly all assumptions met, a uniform distribution of $U(0, 1)$ is expected 
for the p-values. However this is under the assumption that the KS test holds 
its power, which it no longer does due to using fitted parameters. Therefore, 
there is notable deviation from the uniform distribution.

@@

Nonparametric bootstrap can also be used to approximate the null distribution 
of the testing statistic. The procedure is similar, however the resampling is 
performed differently and there is a correction for bias that is required 
\citep{Babu}.
\begin{enumerate}
  \item 
    Draw a random sample $X_1^*,...,X_n^*$ from the empirical distribution $F_n$
    with replacement
  \item 
    Fit $F_\theta$ to the sample and obtain estimated $\hat\theta_n^*$
  \item
    Obtain the empirical distribution function of the random sample $F_n^*$
  \item 
    Calculate bootstrap KS statistic
    \[
      D^* = \sup_x | F_n^* (x)- F_{\hat\theta_n}^*(x) - B_n(x) |.
    \]
    where $B_{n}(x) = \sqrt{n}(F_{n}(x) - F_{\hat\theta_n}(x))$ is the known 
    bias term \citep{Babu}
  \item
    Repeat the previous steps a large number $B$ times and use the empirical 
    distribution of $D^*$ to approximate the null distribution of the observed 
    statistic.
\end{enumerate}
The p-value can be calculated by counting the number of bootstrap KS 
statistics greater than or equal to the observed KS statistic, and then dividing 
by the number of bootstrap samples. Figure~\ref{fig:hist_fitted} displays the 
results of from our simulations. We would expect the distribution of p-values 
to be normal in the case where the KS test holds its power. It is clear from the 
figure that both parametric and nonparametric bootstrap processes correct for 
the problem of fitted parameters. The plots for the bootstrapped p-values appear
to be $U(0,1)$, unlike the naive p-values.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{hist_fitted}
  \caption{The Naive plot (left) is the histogram of p-values from the standard 
  KS test with fitted parameters. The Parametric plot (middle) is the histogram
  of p-values from implementing parametric bootstrap. The Babu plot (right) is 
  the histogram of p-values from implementing the nonparametric bootstrap with a 
  correction for bias. In each case, $1000$ replicate tests were performed on 
  the standard normal distribution with sample size $n = 100$. $B = 1000$ 
  bootstrap samples are obtained for each test.}
  \label{fig:hist_fitted}
\end{figure}

@@

The KS test also displays issues in the case of dependence. As mentioned, an 
assumption of the test is that the data is independent. Unfortunately, real data
is often temporally or spatially dependent and the results of a goodness-of-fit 
test would be valuable. When the KS test is performed on dependent data it 
performs poorly. In this situation, the null hypothesis is $H_0$: the random 
sample $X_1, \ldots, X_n$ come from a continuous distribution $F$ where $F$ has 
completely specified parameters. The test statistic is the same as 
Equation~\ref{eq:ks_standard}. However, since the data is dependent, the null 
hypothesis is wrong and must be corrected. This is demonstrated with a 
simulation in Figure~\ref{fig:hist_correlation}. Data is generated from a 
first-order autoregressive model (AR(1)) with a standard normal distribution. 
P-values are gathered from $1000$ replicate tests for different levels of the 
AR coefficient $\psi$, varying from $(-3,3)$. In the presence of serially 
dependent data, the distribution of p-values no longer follows $U(0, 1)$ as 
would be expected of a valid test. The results echo those of \citet{Durilleul} 
that the KS statistic is too liberal for positive autocorrelation values, and 
that the behavior is asymmetrical for negative values.

@@

Figure~\ref{fig:hist_ar1_D} displays the results of the copula remedy for 
dependent data. The data is generated from the standard normal distribution with
an AR(1) dependence structure where $\psi = .5$. The copula used is to model 
dependence is the normal copula. The distribution of p-values is again expected
to be $U(0, 1)$ for a valid test. The Naive plot is clearly not uniform and 
reinforces the results shown in Figure~\ref{fig:hist_correlation}. The Copula 
plot, which implements the aforementioned procedure to correct for dependence
in the data, appears to be uniform and restores the power of the KS test.

@@

The procedure demonstrated in Section~\ref{sec:dependence} works when the data 
is dependent and the hypothesized distribution is completely specified. However,
this is not practical. In practice we do not know the parameters of the 
hypothesized distribution. Therefore, it is valuable to have a procedure that
corrects for both fitted parameters and serially dependent data. The remedy 
using copulas can be modified to be effective in the case where both assumptions
must be violated. The null hypothesis is equivalent to Section~\ref{sec:fitted} 
and the test statistic is equal to Equation~\ref{eq:ks_fitted}.

@@

Figure~\ref{fig:hist_ar1_FD} shows the results of our simulation on data 
simulated from an AR(1) process. As should be expected, naively fitting 
parameters while providing no adjustment for dependent data invalidates the KS 
test. As well, only using the parametric bootstrap remedy for fitted parameters 
from Section~\ref{sec:fitted} seems to provide weaker results than the 
copula remedy presented above. The results show that copula remedy generates 
uniform p-values and restores the power of the KS test.

@@

Similar to Section~\ref{sec:dependence}, the copula fix is not a complete 
solution. Regardless of the true dependence in the data we assume an AR(1) 
dependence structure by taking the lag-1 sample auto-spearman rho. However, we 
can show that as long as the AR(1) assumption is a close approximation, the 
correction still provides a reasonable approximation that can be useful for 
practical purposes. Figure~\ref{fig:hist_ma1_arma_ar2_FD} shows the results of the 
procedure performed on data generated with dependence structures of  MA(1), 
ARMA(1, 1), and AR(2). Naively fitting parameters and not adjusting for 
dependence clearly deviates from a uniform distribution of p-values. Parametric 
bootstrap provides some correction but does not account for depedence, so 
therefore the results of the copula remedy are more accurate and favorable.
In the case of MA(1) and ARMA(1, 1), the true dependence appears close enough 
to our assumption of AR(1) that the results are reasonable. AR(2) however
appears to be just far enough from our assumption, showing a limitation of the 
procedure.

\begin{figure}[tbp]
  \centering
  \includegraphics{hist_ma1_arma_ar2_FD}
  \caption{The Super Naive plots (left) are the histograms of p-values from the 
  standard KS test with fitted parameters. The Naive Parametric plots (middle) 
  are the histograms of p-values from implementing parametric bootstrap. The 
  Copula plots are the histograms of p-values from implementing parametric 
  bootstrap with copulas for dependence and correcting for fitted parameters. 
  The data is generated from an MA(1) process (left) with $\theta = .5$, ARMA(1,
  1) process (middle) with $\psi = .5$ and $\theta = .3$, and AR(2) process 
  (right) with $\psi = (.5, .3)$ of the standard normal distribution. In each 
  case, $1000$ replicate tests were performed with sample size $n = 100$ and 
  $B = 1000$ bootstrap samples.}
  \label{fig:hist_ma1_arma_ar2_FD}
\end{figure}

@@

\begin{figure}[tbp]
  \centering
  \includegraphics{hist_gamma_gev_FD}
  \caption{The Super Naive plots (left) are the histograms of p-values from the 
  standard KS test with fitted parameters. The Naive Parametric plots (middle) 
  are the histograms of p-values from implementing parametric bootstrap. The 
  Copula plots are the histograms of p-values from implementing parametric 
  bootstrap with copulas for dependence and correcting for fitted parameters. 
  The data is generated from $Gamma(3, 1)$ (left) and $GEV(0, .2, 1)$ (right) 
  with a correlation coefficient of $r = 0.5$. In each case, $1000$ replicate 
  tests were performed with sample size $n = 100$ and $B = 1000$ bootstrap 
  samples.}
  \label{fig:hist_gamma__gev_FD}
\end{figure}

@@

The KS test has base assumptions that the hypothesized distribution is 
completely specified and the data is independent. When these assumptions are 
violated, the test is no longer accurate and remedies must be performed. In the 
case of fitted parameters, parametric and non-parametric bootstrap can restore 
the power of the test. A bias correction is required if the non-parametric form 
is used \citep{Babu}. In the case of dependent data, a procedure using bootstrap 
with copulas to model dependency shows positive results. When both assumptions 
are violated, that is the case where the data has dependence and parameters must 
be fitted, an adjusted copula procedure also shows positive results. The tests 
appear effective for a variety of families of distributions. The copula remedy 
is not a complete solution and has limitations. Regardless of the true 
dependence, we assume an AR(1) depedennce structure. Therefore, if the AR(1) 
dependence structure is a close approximation of the truth, the fix can work. 
However, in cases such as AR(2), if the approximation is too far from the true 
depedence structure the fix does not completely restore power As well, tests 
were only performed with the normal copula. It is possible that other copulas
could provide stronger results based on the true dependence of the data.

>>>>

The Kolmogorov-Smirnov (KS) statistic is one of the most popular goodness-of-fit 
tests for comparing a sample with a hypothesized parametric distribution.
Let $X_1, ..., X_n$ be a random sample from some continuous distribution. 
The null hypothesis $H_0$ is that the population distribution is $F$.
Let $F_n(\cdot)$ be the empirical cumulative distribution. The KS statistic is
\begin{equation}
  \label{eq:ks_standard}
  D = \sup_x | F_{n}(x) - F(x) |.
\end{equation}
The distribution of $D$ under $H_0$ turns out to be independent of the
distribution $F$ and a table of critical values for $D$ has been constructed
\citep{Massey} for various sample sizes $n$ and significance 
levels $\alpha$. If the value of $D$ exceeds the test's corresponding critical 
value the null hypothesis is rejected. The KS test is available in popular 
statistical software packages, such as function \texttt{ks.test} in R 
\citep{R, Marsaglia}.

The standard one-sample KS test applies to independent data with a continuous
hypothesized distribution that is completely specified. Often in literature, it 
has been ``discovered" that the test loses its size when applied to dependent 
or discrete data with hypothesized distributions containing estimated parameters 
\citep{Steinskog}.

@@

The standard KS test is not applicable when the hypothesized distribution 
contains fitted parameters. \citet{Steinskog} ``discovers" the change in power 
when using estimated parameters and stress caution in using the KS test in such 
ways. \citet{Lilliefors} shows for the normal distribution that using the 
standard table when values of the mean and standard deviation are estimated 
obtains extremely conservative results. This is supported by \citet{Capasso}, 
which concludes that failing to re-estimate the parameters may lead to wrong, 
overly-conservative approximations to the distributions of goodness-of-fit test 
statistics based on the empirical distribution function. \citet{Capasso} also 
notes that the impact of this mistake may turn out to be dramatic and does not 
vanish as the sample size increases. Remedies are provided by \citet{Genest} and
\citet{Babu} in the form of bootstrap. \citet{Genest} provides validity for 
using parametric bootstrap with various goodness-of-fit tests. \citet{Babu} 
details the bootstrap procedure for goodness-of-fit tests and notes that both 
parametric and non-parametric procedures lead to correct asymptotic levels, 
however there is a correction required for the non-parametric case.

@@

The rest of the paper is organized as follows. Section~\ref{sec:fitted}
investigates the scenario where the hypothesized distribution has unspecified
parameters. Both parametric and nonparametric bootstrap are available to fix the
issue. Section~\ref{sec:dependence} investigates the scenario where the data of 
the empirical distribution is serially dependent. A bootstrap procedure 
employing copulas is a working solution. Section~\ref{sec:fittedwithdependence}
explores the case where a combination of the first two scenarios occurs. The 
copula procedure can be adjusted for fitted parameters to be available as a fix. 
Section~\ref{sec:conclusion} concludes with a discussion.

@@

To demonstrate the consequences of this problem, a simulation is performed. A 
random sample $X_1, \ldots, X_n$ is generated from the standard normal 
distribution with sample size $n = 100$. The p-values of $1000$ replicate tests 
are displayed in the Naive plot of Figure~\ref{fig:hist_fitted}. Each KS test 
was performed using fitted parameters, i.e. the hypothesized distribution is 
$N(\bar X, s^2)$ where $\bar X$ is the sample mean and $s^2$ is the sample 
variance. Since the data is generated from a standard normal distribution with 
seemingly all assumptions met, a uniform distribution of $U(0, 1)$ is expected 
for the p-values. However this is under the assumption that the KS test holds 
its size, which it no longer does due to using fitted parameters. Therefore, 
there is notable deviation from the uniform distribution.

@@

Nonparametric bootstrap can also be used to approximate the null distribution 
of the testing statistic. The procedure is similar, however the resampling is 
performed differently and there is a correction for bias that is required 
\citep{Babu}.
\begin{enumerate}
  \item 
    Draw a random sample $X_1^*,...,X_n^*$ from the empirical distribution $F_n$
    with replacement
  \item 
    Fit $F_\theta$ to the sample and obtain estimated $\hat\theta_n^*$
  \item
    Obtain the empirical distribution function of the random sample $F_n^*$
  \item 
    Calculate bootstrap KS statistic
    \[
      D^* = \sup_x | F_n^* (x)- F_{\hat\theta_n}^*(x) - B_n(x) |.
    \]
    where $B_{n}(x) = \sqrt{n}(F_{n}(x) - F_{\hat\theta_n}(x))$ is the known 
    bias term \citep{Babu}
  \item
    Repeat the previous steps a large number $B$ times and use the empirical 
    distribution of $D^*$ to approximate the null distribution of the observed 
    statistic.
\end{enumerate}
The p-value can be calculated by counting the number of bootstrap KS 
statistics greater than or equal to the observed KS statistic, and then dividing 
by the number of bootstrap samples. Figure~\ref{fig:hist_fitted} displays the 
results of from our simulations. We would expect the distribution of p-values 
to be uniform in the case where the KS test holds its size. It is clear from the 
figure that both parametric and nonparametric bootstrap processes correct for 
the problem of fitted parameters. The plots for the bootstrapped p-values appear
to be $U(0,1)$, unlike the naive p-values.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{hist_fitted}
  \caption{The Naive plot (left) is the histogram of p-values from the standard 
  KS test with fitted parameters. The Parametric plot (middle) is the histogram
  of p-values from implementing parametric bootstrap. The Babu plot (right) is 
  the histogram of p-values from implementing nonparametric bootstrap with a 
  correction for bias. In each case, $1000$ replicate tests were performed on 
  the standard normal distribution with sample size $n = 100$. $B = 1000$ 
  bootstrap samples are obtained for each test.}
  \label{fig:hist_fitted}
\end{figure}

@@

The KS test also displays issues in the case of dependence. As mentioned, an 
assumption of the test is that the data is independent. Unfortunately, real data
is often temporally or spatially dependent and the results of a goodness-of-fit 
test would be valuable. When the KS test is performed on dependent data it 
performs poorly. In this situation, the null hypothesis is $H_0$: the random 
sample $X_1, \ldots, X_n$ come from a continuous distribution $F$ where $F$ has 
completely specified parameters. The test statistic is the same as 
Equation~\eqref{eq:ks_standard}. However, since the data is dependent, the null 
hypothesis is wrong and must be corrected. This is demonstrated with a 
simulation in Figure~\ref{fig:hist_correlation}. Data is generated from a 
first-order autoregressive model (AR(1)) with a standard normal distribution. 
P-values are gathered from $1000$ replicate tests for different levels of the 
AR coefficient $\psi$, varying from $(-3,3)$. In the presence of serially 
dependent data, the distribution of p-values no longer follows $U(0, 1)$ as 
would be expected of a valid test. The results echo those of \citet{Durilleul} 
that the KS statistic is too liberal for positive autocorrelation values, and 
that the behavior is asymmetrical for negative values.

@@

Figure~\ref{fig:hist_ar1_D} displays the results of the copula remedy for 
dependent data. The data is generated from the standard normal distribution with
an AR(1) dependence structure where $\psi = .5$. The copula used is to model 
dependence is the normal copula. The distribution of p-values is again expected
to be $U(0, 1)$ for a valid test. The Naive plot is clearly not uniform and 
reinforces the results shown in Figure~\ref{fig:hist_correlation}. The Copula 
plot, which implements the aforementioned procedure to correct for dependence
in the data, appears to be uniform and restores the size of the KS test.

@@

The procedure demonstrated in Section~\ref{sec:dependence} works when the data 
is dependent and the hypothesized distribution is completely specified. However,
this is not practical. In practice we do not know the parameters of the 
hypothesized distribution. Therefore, it is valuable to have a procedure that
corrects for both fitted parameters and serially dependent data. The remedy 
using copulas can be modified to be effective in the case where both assumptions
must be violated. The null hypothesis is equivalent to Section~\ref{sec:fitted} 
and the test statistic is equal to Equation~\eqref{eq:ks_fitted}.

@@

Figure~\ref{fig:hist_ar1_FD} shows the results of our simulation on data 
simulated from an AR(1) process. As should be expected, naively fitting 
parameters while providing no adjustment for dependent data invalidates the KS 
test. As well, only using the parametric bootstrap remedy for fitted parameters 
from Section~\ref{sec:fitted} seems to provide weaker results than the 
copula remedy presented above. The results show that copula remedy generates 
uniform p-values and restores the size of the KS test.

@@

Similar to Section~\ref{sec:dependence}, the copula fix is not a complete 
solution. Regardless of the true dependence in the data we assume an AR(1) 
dependence structure by taking the sample first-order Spearman's rho. However, 
we can show that as long as the AR(1) assumption is a close approximation, the 
correction still provides a reasonable results that can be useful for 
practical purposes. Figure~\ref{fig:hist_ma1_arma_ar2_FD} shows the results of 
the procedure performed on data generated with dependence structures MA(1), 
ARMA(1, 1), and AR(2). Naively fitting parameters and not adjusting for 
dependence clearly deviates from a uniform distribution of p-values. Parametric 
bootstrap provides some correction but does not account for depedence, so 
therefore the results of the copula remedy are more accurate and favorable.
In the case of MA(1) and ARMA(1, 1), the true dependence appears close enough 
to our assumption of AR(1) that the results are reasonable. AR(2) however
appears to be just far enough from our assumption, showing a limitation of the 
procedure.

\begin{figure}[tbp]
  \centering
  \includegraphics{hist_ma1_arma_ar2_FD}
  \caption{The Super Naive plots (top) are the histograms of p-values from the 
  standard KS test with fitted parameters. The Naive Parametric plots (middle) 
  are the histograms of p-values from implementing parametric bootstrap. The 
  Copula plots (bottom) are the histograms of p-values from implementing 
  parametric bootstrap with copulas for dependence and correcting for fitted 
  parameters. The data is generated from an MA(1) process (left) with 
  $\theta = .5$, ARMA(1, 1) process (middle) with $\psi = .5$ and $\theta = .3$,
  and AR(2) process (right) with $\psi = (.5, .3)$ of the standard normal 
  distribution. In each case, $1000$ replicate tests were performed with sample 
  size $n = 100$ and $B = 1000$ bootstrap samples.}
  \label{fig:hist_ma1_arma_ar2_FD}
\end{figure}

@@

\begin{figure}[tbp]
  \centering
  \includegraphics{hist_gamma_gev_FD}
  \caption{The Super Naive plots (left) are the histograms of p-values from the 
  standard KS test with fitted parameters. The Naive Parametric plots (middle) 
  are the histograms of p-values from implementing parametric bootstrap. The 
  Copula plots are the histograms of p-values from implementing parametric 
  bootstrap with copulas for dependence and correcting for fitted parameters. 
  The data is generated from $Gamma(3, 1)$ (left) and $GEV(0, .2, 1)$ (right) 
  with a correlation coefficient of $r = 0.5$. In each case, $1000$ replicate 
  tests were performed with sample size $n = 100$ and $B = 1000$ bootstrap 
  samples.}
  \label{fig:hist_gamma_gev_FD}
\end{figure}

@@

The KS test has base assumptions that the hypothesized distribution is 
completely specified and the data is independent. When these assumptions are 
violated, the test is no longer accurate and remedies must be performed. In the 
case of fitted parameters, parametric and non-parametric bootstrap can restore 
the size of the test. A bias correction is required if the non-parametric form 
is used \citep{Babu}. In the case of dependent data, a procedure using bootstrap 
with copulas to model dependency shows positive results. When both assumptions 
are violated, that is the case where the data has dependence and parameters must 
be fitted, an adjusted copula procedure also shows positive results. The tests 
appear effective for a variety of families of distributions. The copula remedy 
is not a complete solution and has limitations. Regardless of the true 
dependence, we assume an AR(1) depedence structure. Therefore, if the AR(1) 
dependence structure is a close approximation of the truth, the fix can work. 
However, in cases such as AR(2), if the approximation is too far from the true 
depedence structure the fix does not completely restore size As well, tests 
were only performed with the normal copula. It is possible that other copulas
could provide stronger results based on the true dependence of the data.
